Reduced MHD Manuscript

### 2024-12-29 Manuscript/Manuscript.tex
Edits

1: MHD Eqns. 2: Split para/perp. Alg from two. 3: My nd process. 4: Improving Strauss nd process. 5: Fast-Slow Attempt. 6: 2D Decomps and q. 7: Take 4 and change vars. 8: New wFSS. Appendix: Lemmas and Identities.

@@

\section{Nondimensionalization and Ordering} 
The dependent variables in our system of PDEs are highly coupled to one another. We will give a formal description of these relationships by nondimensionalizing our MHD model and comparing the dimensionless parameters that remain. These parameters will define an ordering/perturbative expansion, an observation timescale, and will characterize different dynamical regimes.

Because our domain factorizes, it is convenient to nondimensionalize our coordinates according to their characteristic length scales: 
\begin{align*}
    x=aX,\quad y=aY,\quad z=\frac{L}{2\pi}Z.
\end{align*}
This induces a separation between derivatives in the poloidal and toroidal directions, $\nabla = \frac{1}{a}\np + \frac{2\pi}{L}\bm{e}_Z\p_Z$, where $\np = \pth{\p_X, \p_Y, 0}$. For example, the divergence condition becomes   
\begin{align*}
    \nabla\cdot\bm{B} = \frac{1}{a}\np\cdot\bm{B}_\perp + \frac{2\pi}{L}\p_ZB_Z = 0. 
\end{align*}

Many plasma experiments take place on large aspect-ratio toroidal domains, where the ratio $\ep = 2\pi a/L$ is small. In such experiments, the magnetic field is usually much stronger in the toroidal than the poloidal directions, and the variations are small. This behavior motivates us to choose an ordering that splits $\bm{B}$ accordingly: 
\begin{align*}
    \rho &= \rho_0 r \\
    \bm{v} &= v_0 \bm{\nu} \\ 
    \bm{B} &= B_0 \pmat{\ep\beta_x \\ \ep\beta_y \\ 1+\ep^2\beta_\para}. 
\end{align*}

****

The only quantities in our system that still have units are time and pressure. We choose $p=p_0\pi(r)$, where , we are able to identify the following parameters as they appear in our nondimensionalized MHD system:  
\begin{align*}
    \frac{t_0 v_0}{a} = 1, \qquad 
    \frac{p_0}{\mu_0^{-1}B_0^2} = \beta_0, \qquad 
    \frac{\rho_0\,v_0^2}{\mu_0^{-1}B_0^2} = M_0^2\beta_0. 
\end{align*}
The first of these parameters establishes a particular observation timescale, where we are zoomed in on dynamics in the poloidal plane. The second and third define the plasma-beta parameter, $\beta_0$, and the Mach number, $M_0$. respectively. Different orderings of these quantities in $\ep$ correspond to different regimes, where plasma behavior is characteristically different.

\subsection{Rescaled Continuity Equation}
\begin{align*}
    \frac{\rho_0}{t_0}\deriv{r}{\tau} &= -\frac{\rho_0v_0}{a}\np\cdot\pth{r\bm{\nu}_\perp} - \frac{2\pi\rho_0v_0}{L} \p_Z\pth{r\nu_\para}
\end{align*}
or

\subsection{Rescaled Momentum Conservation}

\subsection{Rescaled Faraday Induction}
\begin{align*} 
\deriv{B_z}{t} &= 
    \pth{B_\perp\cdot\np + B_z\p_z}v_z - \br{(v_\perp\cdot\np + v_z\p_z) + (\np\cdot v_\perp + \p_zv_z)} B_z \\
    &= B_\perp\cdot\np v_z - v_\perp\cdot\np B_z - v_z\p_zB_z - \np \cdot v_\perp B_z 
\end{align*}

\section{Improving Strauss's Scaling}
$\rho$ motivation: change $r$ because fluctuations in $\rho$, $\p_\tau \rho$ should be order $\ep$, not order 1. Earlier scaling overestimated importance of $r$ in evolution, which was not physical. This is the proper way of incorporating order $\ep$ changes, because higher order terms go away in limit.

B motivation: $\beta_\para$ fluctuates more than we originally put. 
\begin{align*}
    \rho &= \rho_0\,(1 + \epsilon\,r)\\
    \bm{v} & = v_0\,\bm{\nu} = v_0 \pmat{} \\
    \bm{B} & = B_0\bigg((1 + \epsilon\,\beta_\parallel)\bm{e}_z + \epsilon\,\bm{\beta}_\perp\bigg) = B_0 \pmat{\ep\beta_x \\ \ep\beta_y \\ 1+\ep\beta_\para} .
\end{align*}

>>>>

1: MHD Eqns. 2: ND process and ordering. 3: Split para/perp. Alg from two. 4: Fast-Slow Attempt. 5: 2D Decomps and q. 6: Take 4 and change vars. 7: New wFSS. Appendix: Lemmas and Identities.

@@

\section{Nondimensionalization and Ordering} 
The dependent variables in our system of PDEs are highly coupled to one another. We will give a formal description of these relationships by nondimensionalizing our MHD model and comparing the dimensionless parameters that remain. These parameters will characterize an ordering, an observation timescale, and various dynamical regimes.

Because our domain factorizes, it is convenient to nondimensionalize our coordinates according to their characteristic length scales: 
\begin{align*}
    x=aX,\quad y=aY,\quad z=\frac{L}{2\pi}Z.
\end{align*}
This allows us to replace the explicit dependence on both $a$ and $L$ with just the aspect ratio $\ep = 2\pi a/L$. For example, if we multiply each equation by a factor of $a$, the standard gradient becomes $a\nabla = \np + \ep\bm{e}_Z\p_Z$, where $\np = \pth{\p_X, \p_Y, 0}$. In these coordinates, the divergenceless condition is    
\begin{align*}
    a\nabla\cdot\bm{B} = \np\cdot\bm{B}_\perp + \ep\p_Z B_Z = 0, \quad\text{or}\quad \np\cdot\bm{B}_\perp = -\ep\p_ZB_Z. 
\end{align*}

Many plasma experiments take place on large aspect-ratio toroidal domains, where the ratio $\ep = 2\pi a/L$ is small. In such experiments, the magnetic field is usually much stronger in the toroidal than the poloidal directions, and the variations are small. Density is also relatively constant when $\ep\to 0$. This behavior motivates us to examine the specific ordering where 
\begin{align*}
    \rho &= \rho_0 \pth{1+\ep r} \\
    \bm{v} &= v_0 \bm{\nu} \\ 
    \bm{B} &= B_0 \pmat{\ep\beta_x \\ \ep\beta_y \\ 1+\ep\beta_\para},
\end{align*}
with $r, \bm{\nu}, \beta_x, \beta_y,$ and $\beta_\para$ all $O\pth{1}$ in $\ep$. The constant coefficients $\rho_0, v_0,$ and $B_0$ represent values typical of a given experiment, which can be relatiely small or large. Time and pressure will be treated similarly, so that $t=t_0\tau$ and $p=p_0\pi(r)$ with $\tau,\pi(r) = O(1)$.

This scaling is similar to one taken by Strauss in 1975. In his scaling however, toroidal magnetic field fluctuations are assumed to be $O\pth{\ep^2}$, restricting the kinds of scenarios where this process can be applied. Despite density fluctuations being small for small $\ep$, he also allows variations at $O(1)$, so that 
\begin{align*}
    \rho &= \rho_0 r \\
    \bm{v} &= v_0 \bm{\nu} \\ 
    \bm{B} &= B_0 \pmat{\ep\beta_x \\ \ep\beta_y \\ 1+\ep^2\beta_\para}.
\end{align*}
These assumptions are updated in the current analysis.

When we rewrite our nondimensionalized MHD system in the following sections, we will find the following dimensionless ratios:  
\begin{align*}
    \frac{t_0 v_0}{a} = 1, \qquad 
    \frac{p_0}{\mu_0^{-1}B_0^2} = \beta_0, \qquad 
    \frac{\rho_0\,v_0^2}{p_0} = M_0^2. 
\end{align*}
The first of these establishes a particular observation timescale, where we are zoomed in on dynamics in the poloidal plane. The second and third define the plasma-beta parameter, $\beta_0$, and Mach number, $M_0$, respectively. Different orderings of these quantities in $\ep$ correspond to different regimes, where plasma behavior is characteristically different.

\section{Rescaled Evolution Equations}
\subsection{Continuity Equation}
In our new dimensionless coordinates, the continuity equation reads
\begin{align*}
    a\deriv{\rho}{t} &= -\np\cdot\pth{\rho\bm{v}_\perp} - \ep\p_Z\pth{\rho v_Z}. 
\end{align*}
\begin{align*}
    \ep\frac{a\rho_0}{t_0}\deriv{r}{\tau} &= -a\nabla\cdot\pth{\rho\bm{v}} \\ 
    &= -\rho_0v_0\np\cdot\pth{\pth{1+\ep r}\bm{\nu}_\perp} - \ep\rho_0v_0 \p_Z\pth{\pth{1+\ep r}\nu_\para} \\ 
    &= -\rho_0v_0 a\nabla\cdot\pth{\pth{1+\ep r} \bm{\nu}}, 
\end{align*}
or, identifying our scaling coefficients, 
\begin{align*} 
    \ep\deriv{r}{\tau} &= -\frac{t_0v_0}{a} a\nabla\cdot\pth{\pth{1+\ep r}\bm{\nu}}
    = -\np\cdot\pth{\pth{1+\ep r}\bm{\nu}_\perp} - \ep\p_Z\pth{\pth{1+\ep r}\nu_\para}. 
\end{align*}

\subsection{Momentum Conservation}
Equation \eqref{momentum} contains three terms which must be separated into their poloidal and toroidal components in order to identify dimensionless parameters.
\begin{align*}
    \frac{a\rho_0v_0}{t_0} \pth{1+\ep r} \deriv{\bm{\nu}}{\tau} &= \mu_0^{-1} \pth{a\nabla\times \bm{B}} \times \bm{B} - a\nabla p(\rho) - \rho \bm{v}\cdot a\nabla \bm{v}. 
\end{align*}
To isolate the evolution of $\bm{\nu}$, we will remove $\frac{\rho_0v_0}{t_0} \pth{1+\ep r}$ from both sides.

For the first term, we make use of the identity $\pth{\nabla\times\bm{B}}\times\bm{B} = \bm{B}\cdot\nabla\bm{B} - \pth{\nabla\bm{B}}\cdot\bm{B}$ to get 
\begin{align*}
    \br{\pth{a\nabla\times\bm{B}}\times\bm{B}}_Z &= \bm{B}\cdot a\nabla B_Z - \ep\pth{\p_Z\bm{B}}\cdot\bm{B} \\ 
    &= \pth{\bm{B}_\perp\cdot\np B_Z + \ep B_Z\p_Z B_Z} - \pth{\ep\bm{B}_\perp\cdot\p_Z\bm{B}_\perp + \ep B_Z\p_ZB_Z} \\ 
    &= \bm{B}_\perp\cdot \np B_Z + \ep\bm{B}_\perp\cdot \p_Z\bm{B}_\perp \\ 
    &= B_0^2 \pth{\ep^2\bm{\beta}_\perp\cdot\np\beta_\para + \ep^3 \bm{\beta}_\perp\cdot\p_Z\bm{\beta}_\perp}  \quad\text{and}
\end{align*}
Note! $\beta_\perp\cdot\p_Z\beta_\perp = \frac{1}{2}\p_Z|\beta_\perp|^2$. Which expression is handier? 
\begin{align*}
    \br{\pth{a\nabla\times\bm{B}}\times\bm{B}}_\perp &= \bm{B}\cdot a\nabla\bm{B}_\perp - \pth{\np\bm{B}}\cdot\bm{B} \\ 
    &= \pth{\bm{B}_\perp\cdot\np\bm{B}_\perp + \ep B_Z\p_Z\bm{B}_\perp} - \pth{\pth{\np\bm{B}_\perp} \cdot\bm{B}_\perp + B_Z\np B_Z} \\ 
    &= \pth{\np\times\bm{B}_\perp}\times\bm{B}_\perp + B_Z\pth{\ep\p_Z\bm{B}_\perp - \np B_Z} \\ 
    &= \ep^2 B_0^2 \pth{\np\times\bm{\beta}_\perp}\times\bm{\beta}_\perp + B_0^2 \pth{1+\ep\beta_\para} \pth{\ep^2\p_Z\bm{\beta}_\perp - \ep\np\beta_\para}. 
\end{align*}

The pressure term gives 
\begin{align*}
    \nabla p\pth{\rho} &= \frac{p_0}{a}\np\pi\pth{1+\ep r} + \bm{e}_Z \frac{2\pi p_0}{L}\p_Z\pi.
\end{align*}

\begin{align*}
    \frac{a}{t_0v_0}\frac{\rho_0v_0^2}{\mu_0^{-1}B_0^2} \pth{1+\ep r} \deriv{\bm{\nu}}{\tau} = M_0^2\beta_0 \pth{1+\ep r} \deriv{\bm{\nu}}{\tau} &= \frac{1}{B_0^2} \pth{a\nabla\times\bm{B}}\times\bm{B} - \beta_0 a\nabla\pi - M_0^2\beta_0 \bm{\nu}\cdot a\nabla\bm{\nu}. 
\end{align*}
where we've multiplied through by $a/\mu_0^{-1}B_0^2$ to make use of orderings. $a\nabla = \np + \ep\bm{e}_Z\p_Z$ makes things easier too.

\subsection{Faraday Induction}
\begin{align*} 
\deriv{B_z}{t} &= 
    \pth{B_\perp\cdot\np + B_z\p_z}v_z - \br{(v_\perp\cdot\np + v_z\p_z) + (\np\cdot v_\perp + \p_zv_z)} B_z \\
    &= B_\perp\cdot\np v_z - v_\perp\cdot\np B_z - v_z\p_zB_z - \np \cdot v_\perp B_z 
\end{align*}
